---
title: "CRU climate forcing ingest for running rsofun"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
---

```{r setup, include=FALSE}
library(tidyverse)
library(ingestr)
library(rbeni)
library(rsofun)
```

## Meta info

```{r}
df_sites <- siteinfo_fluxnet2015 %>% 
  dplyr::filter(sitename %in% c("CH-Lae", "CH-Dav")) %>% 
  mutate(year_start = 1901, year_end = 2018)
```


## Climate forcing

See vignette in ingestr for testing against station data CH-Lae.

Get CRU data and bias-correct. Bias correction based on high-resolution WorldClim 1970-2000 monthly climatology is available for variables temp, prec, and vpd.
```{r message=FALSE}
## get daily data (with temporal downscaling)
df_forcing_cru <- ingest(
  df_sites,
  source    = "cru",
  getvars   = c("temp", "prec", "vpd", "ccov"),
  dir       = "~/data/cru/ts_4.05/",
  timescale = "d",
  settings  = list(correct_bias = "worldclim", dir_bias = "~/data/worldclim")
  )

# df_forcing_cru$data[[1]] %>% 
#   ggplot(aes(date, vpd)) +
#   geom_line()
```

Get PPFD from WATCH-WFDEI and assume a means seasonal cycle for year before 1979.
```{r}
df_forcing_watch <- ingest(
  df_sites %>% mutate(year_start = 1979),
  source    = "watch_wfdei",
  getvars   = c("ppfd"),
  dir       = "~/data/watch_wfdei/",
  timescale = "d",
  settings  = list(correct_bias = "worldclim", dir_bias = "~/data/worldclim")
  )
```

Combine. For pre-1979 years, take mean seasonal cycle for PPFD. For all years, use means seasonal cycle for CCOV (cloud cover fraction).
```{r}
filn <- "data/df_forcing.rds"
if (!file.exists(filn)){
  df_msc_watch <- df_forcing_watch %>% 
    unnest(data) %>% 
    mutate(doy = yday(date)) %>% 
    group_by(sitename, doy) %>% 
    summarise(ppfd_msc = mean(ppfd, na.rm = TRUE))

  df_msc_cru <- df_forcing_cru %>% 
    unnest(data) %>% 
    mutate(doy = yday(date)) %>% 
    group_by(sitename, doy) %>% 
    summarise(ccov_msc = mean(ccov, na.rm = TRUE))
  
  df_forcing <- df_forcing_cru %>% 
    unnest(data) %>% 
    left_join(df_forcing_watch %>% unnest(data) %>% dplyr::select(sitename, date, ppfd), by = c("sitename", "date")) %>% 
    mutate(doy = yday(date)) %>% 
    left_join(df_msc_watch, by = c("sitename", "doy")) %>% 
    left_join(df_msc_cru, by = c("sitename", "doy")) %>% 
    mutate(ppfd = ifelse(is.na(ppfd), ppfd_msc, ppfd)) %>% 
    
    ## add standard atmospheric pressure corrected by elevation
    left_join(df_sites %>% dplyr::select(sitename, elv), by = "sitename") %>% 
    mutate(patm = calc_patm(elv = elv)) %>%
    
    ## assume precip is snow when temp is below 1 deg
    mutate(snow = ifelse(temp < 1, prec, 0),
           rain = ifelse(temp >= 1, prec, 0)) %>% 
    
    ## overwrite cloud cover with mean seasonal cycle
    mutate(ccov = ccov_msc)
  
  saveRDS(df_forcing, file = filn)
    
} else {
  df_forcing <- readRDS(filn)
}
```


### fAPAR forcing

```{r}
df_fapar <- ingest(
  siteinfo  = df_sites,
  source    = "fapar_unity"
  )
```

### CO2 forcing

Ingesting CO2 data is particularly simple. We can safely assume it's well mixed in the atmosphere (independent of site location), and we can use a annual mean value for all days in respective years.

```{r}
df_co2 <- ingestr::ingest(
  df_sites,
  source  = "co2_cmip",
  dir = "~/data/co2"
  )
```

## Simulation settings

```{r}
params_siml <- list(
  spinup             = TRUE,      # to bring soil moisture to steady state
  spinupyears        = 10,        # number of spinup years. 10 is enough for soil moisture.
  recycle            = 1,         # number of years recycled during spinup 
  soilmstress        = FALSE,     # boolean for whether soil moisture stress function is included
  tempstress         = TRUE,      # boolean for whether temperature stress function is included
  calc_aet_fapar_vpd = FALSE,     # set to FALSE - should be dropped again
  in_ppfd            = TRUE,      # if available from forcing files, set to TRUE
  in_netrad          = FALSE,     # if available from forcing files, set to TRUE
  outdt              = 1,
  ltre               = FALSE,
  ltne               = FALSE,
  ltrd               = FALSE,
  ltnd               = FALSE,
  lgr3               = TRUE,
  lgn3               = FALSE,
  lgr4               = FALSE
  )
```

<!-- Run `prepare_setup_sofun()` to define the simulation settings that contain all the information specified by the two steps above (meta info, and simulation parameters), global simulation parameters are wrapped inside an additional column `params_siml`, added to the site meta info dataframe. -->

<!-- ```{r} -->

<!-- siteinfo <- prepare_setup_sofun(siteinfo = siteinfo, params_siml = params_siml) -->

<!-- ``` -->

## Model parameters

```{r}
params_modl <- list(
  kphio           = 0.09423773,
  soilm_par_a     = 0.33349283,
  soilm_par_b     = 1.45602286
  )
```

## Soil parameters

```{r}
df_soiltexture <- bind_rows(
  top    = tibble(layer = "top",    fsand = 0.4, fclay = 0.3, forg = 0.1, fgravel = 0.1),
  bottom = tibble(layer = "bottom", fsand = 0.4, fclay = 0.3, forg = 0.1, fgravel = 0.1)
)
```

## Collect all drivers

```{r}
df_drivers <- collect_drivers_sofun( 
  siteinfo       = df_sites,
  params_siml    = params_siml,
  meteo          = df_forcing %>% group_by(sitename) %>% nest(), 
  fapar          = df_fapar,
  co2            = df_co2,
  df_soiltexture = df_soiltexture
  )
df_drivers
```

## Run model

```{r}
df_output <- runread_pmodel_f(
     df_drivers,
     params_modl = params_modl, 
     makecheck = TRUE,
     parallel = FALSE
     ) %>% 
  unnest(data)

saveRDS(df_output, file = "data/df_output.rds")
```


## Process and visualise

```{r}
df_output %>% 
  mutate(year = year(date), chi_wgt = chi * gpp) %>% 
  group_by(sitename, year) %>% 
  summarise(chi_wgt = sum(chi_wgt), gpp = sum(gpp)) %>% 
  mutate(chi = chi_wgt / gpp) %>% 
  ggplot(aes(x = year, y = chi, color = sitename)) +
  geom_line()

df_output %>% 
  mutate(year = year(date), iwue_wgt = iwue * gpp) %>% 
  group_by(sitename, year) %>% 
  summarise(iwue_wgt = sum(iwue_wgt), gpp = sum(gpp)) %>% 
  mutate(iwue = iwue_wgt / gpp) %>% 
  ggplot(aes(x = year, y = iwue, color = sitename)) +
  geom_line()
```


